Do Alternate Wetting and Drying Irrigation Technology and Nitrogen Rates Affect Rice Sheath Blight?

### 2020-07-09 analysis/paper/paper.Rmd
Flesh out stats methods and results, update figure files

theme_set(theme_classic(base_size = 14))

@@

Bayesian multivariate mixed models were fitted using the contributed R package `MCMCglmm`, version 2.29, [@Hadfield2010].
Six models were created where the dependent variables were: 2015 tiller sheath blight incidence, 2015 tiller sheath blight severity, 2015 leaf sheath blight severity; 2016 tiller sheath blight incidence, 2016 tiller sheath blight severity; 2016 leaf sheath blight severity, respectively.
The independent variables were the irrigation management (WMGT) and N rate (NRTE) and the interaction of WMGT and NRTE with replicate treated as a random variable.
The base levels (control treatments) for the analyses for 2015 were: N rate -- N0, irrigation management -- puddled. For for 2016 the base level values were: N rate -- N60, irrigation management -- puddled.

The same weakly informative priors were set for the fixed and random effects in all models, $V = 1$, $nu = 0.02$.
Models were run for 60,000 iterations with every ten being selected.
Model fit was evaluated using `plot.MCMCglmm()` and `summary.MCMCglmm()`.

@@

In 2015 the incidence of tiller sheath blight remained low throughout the growing season (Fig.&nbsp;\@ref(fig:fig-1)a, \@ref(fig:fig-2)a).
Water management was not clearly different (Fig.&nbsp;\@ref(fig:fig-3)a, \@ref(fig:fig-5)a).
However, the effect of the N treatments N120 on tiller incidence was higher and clearly different when compared with the control N0 treatment (Fig.&nbsp;\@ref(fig:fig-5)a, \@ref(fig:fig-5)a).

Tiller sheath blight severity remained below 2&nbsp;\% (Fig.&nbsp;\@ref(fig:fig-2)c, \@ref(fig:fig-1)c).
Both the N100 and N120 treatment effects caused the tiller severity to be higher and clearly statistically different than the control N0 treatment (Fig.&nbsp;\@ref(fig:fig-4)c, \@ref(fig:fig-5)c).
However, the AWD water management was not clearly statistically different from the puddled treatment (Fig.&nbsp;\@ref(fig:fig-3)c, \@ref(fig:fig-6)a).

The severity of leaf sheath blight remained low, less than 0.4&nbsp;% across all treatments (Fig.&nbsp;\@ref(fig:fig-1)e, \@ref(fig:fig-2)e).
As with the tiller incidence and severity, the effects of both the N100 and N120 treatments on leaf severity resulted in higher values and were and clearly statistically different than the N0 treatment (Fig.&nbsp;\@ref(fig:fig-4)e, \@ref(fig:fig-5)e).
The AWD water management was clearly lower than the puddled treatment (Fig.&nbsp;\@ref(fig:fig-3)e, \@ref(fig:fig-7)a).

The interaction of irrigation treatment and nitrogen fertilisation was not clearly different for any of the three response variables analysed.

@@

The changes to the inoculation methodology resulted in a higher rate of infection in 2016, with the N180 treatment reaching a maximum value of 98&nbsp;% incidence at the third observation (Fig.&nbsp;\@ref(fig:fig-2)b).
The N treatment N180 value was higher and clearly statistically different than the control N60 treatment (Fig.&nbsp;\@ref(fig:fig-4)b, \@ref(fig:fig-5)b). As in the 2015 study, water management did not clearly statistically differ (Fig.&nbsp;\@ref(fig:fig-3)b, \@ref(fig:fig-5)b).

As with the tiller incidence, the tiller severity increased with the changed inoculation methods with a maximum of 7.6% for the puddled treatment (Fig.&nbsp;\@ref(fig:fig-2)d).
In 2016 the N180 treatment was again higher and clearly statistically different from the N60 treatment, with N180 severity being higher (Fig.&nbsp;\@ref(fig:fig-4)d, \@ref(fig:fig-5)d).
The AWD water management, which was lower than the puddled treatment, was also clearly statistically different (Fig.&nbsp;\@ref(fig:fig-3)d, \@ref(fig:fig-6)b).

In 2016 the neither of the treatments, N rate or water management, were clearly statistically different from the control treatments, N60 and puddling water management, for leaf sheath blight severity (Fig.&nbsp;\@ref(fig:fig-3)f, \@ref(fig:fig-5)b, \@ref(fig:fig-4)b, \@ref(fig:fig-7)b).

The interaction of irrigation treatment AWD with N rate of N180 in 2016 was not clearly different from the control puddled irrigation and N60 for tiller incidence or severity in either experiment.
In the 2016 experiment, the interactions were not clearly different for leaf severity from the control puddled irrigation and N60.
In 2016 the interaction of N180 and AWD was clearly lower than the control puddled irrigation and N60 rate.

@@

x <- flextable(x)
x <- fontsize(x, part = "header", size = 10)
x <- font(x, part = "header", fontname = "Times")
x <- font(x, fontname = "Times")
x <- fontsize(x, size = 10)
x <- set_caption(x, 
  "(\\#tab:Table 2) An 89 % test for practical equivalence for 2015 experiment. Response variables are shown as T. Inc., Tiller Incidence; T. Sev., Tiller Severity; L. Sev., Leaf Severity. The parameters are WMGTAWD, irrigation treatment AWD; NRTEN100, nitrogen rate of 100 k/ha; NRTEN120 kg/ha. ROPE low and ROPE high provide the limits of the Region of Practical Equivalence (ROPE). ROPE Percentage is the proportion of the Highest Density Interval (HDI) that lies inside the ROPE. All points within this interval have a higher probability of density than points outside the interval. ROPE Equivalence is the test result, either \"rejected\", \"accepted\" or \"undecided\". \"HDI Low\" and \"HDI High\" are the lower and upper HDI limits for the parameters."
  )
x
```

@@

x <- flextable(x)
x <- fontsize(x, part = "header", size = 10)
x <- font(x, part = "header", fontname = "Times")
x <- font(x, fontname = "Times")
x <- fontsize(x, size = 10)
x <- set_caption(x, 
  "(\\#tab:Table 2) An 89 % test for practical equivalence for 2015 experiment. Response variables are shown as T. Inc., Tiller Incidence; T. Sev., Tiller Severity; L. Sev., Leaf Severity. The parameters are WMGTAWD, irrigation treatment AWD; and NRTEN180, nitrogen rate of 180 k/ha. ROPE low and ROPE high provide the limits of the Region of Practical Equivalence (ROPE). ROPE Percentage is the proportion of the Highest Density Interval (HDI) that lies inside the ROPE. All points within this interval have a higher probability of density than points outside the interval. ROPE Equivalence is the test result, either \"rejected\", \"accepted\" or \"undecided\". \"HDI Low\" and \"HDI High\" are the lower and upper HDI limits for the parameters.")
x
```

>>>>

theme_set(theme_classic(base_size = 8))

@@

The contributed R package, rstanarm (version 2.19.3), [@Brilleman2018; @Goodrich2020], was used to fit Bayesian generalized linear models with water management and nitrogen fertilisation as fixed effects and replicate as a random effect.
Six models were created where the dependent variables were: 2015 tiller sheath blight incidence, 2015 tiller sheath blight severity, 2015 leaf sheath blight severity; 2016 tiller sheath blight incidence, 2016 tiller sheath blight severity; 2016 leaf sheath blight severity, respectively.
The independent variables were the irrigation management (WMGT) and N rate (NRTE) and the interaction of WMGT and NRTE with replicate treated as a random variable.
The base levels (control treatments) for the analyses for 2015 were: N rate -- N0, irrigation management -- puddled. For for 2016 the base level values were: N rate -- N60, irrigation management -- puddled.
The same weakly informative priors were set for the fixed and random effects in all models, location = 0, scale = 10.
Model fitness was evaluated using `mcmc_trace()` and `pp_check()` from bayesplot, version 1.7.2, [@Gabry2019] as well as summary information.

The resulting models were tested using a test for equivalency using `equivalence_test()` from bayestestR, version 0.7.0, [@Makowski2019] to evaluate whether differences between the sheath blight responses varied between the water and nitrogen management treatments (Tab.&nbsp;\@ref(tab:Table-2), Tab.&nbsp;(tab:Table-3)). The Test for Practical Equivalence is based on the Highest Density Interval (HDI) plus the Region of Practical Equivalence (ROPE) decision rule, *HDI+ROPE decision rule* and checks whether parameter values should be accepted or rejected against the null hypothesis, the ROPE [@Kruschke2018]. To do this, it checks the percentage of the 89&nbsp;% HDI that is in the null region ROPE. If this amount is sufficiently low, the null hypothesis is rejected. If the amount is large enough, the null hypothesis is accepted.

## Hypothesis {-}

We hypothesised that higher amounts of nitrogen would increase the incidence and severity of sheath blight, while alternate wetting and drying would decrease the disease incidence and severity.

@@

In 2015 the incidence of tiller sheath blight remained low throughout the growing season (Fig.&nbsp;\@ref(fig:fig-1)a, \@ref(fig:fig-2)a).
Water management was undecided in the test for equivalency (Fig.&nbsp;\@ref(fig:fig-3)a, \@ref(fig:fig-5)a) between the AWD and puddled treatments.
However, the effect of the N treatments of both N100 and N120 on tiller incidence were higher and clearly different when compared with the control N0 treatment (Fig.&nbsp;\@ref(fig:fig-5)a, \@ref(fig:fig-5)a).

Tiller sheath blight severity remained below 2&nbsp;\% (Fig.&nbsp;\@ref(fig:fig-2)c, \@ref(fig:fig-1)c).
Both the N100 and N120 treatment effects caused the tiller severity to be higher and were clearly different than the control N0 treatment (Fig.&nbsp;\@ref(fig:fig-4)c, \@ref(fig:fig-5)c).
However, the AWD water management was not clearly the same or different from the puddled treatment (Fig.&nbsp;\@ref(fig:fig-3)c, \@ref(fig:fig-6)a).

The severity of leaf sheath blight remained low, less than 0.4&nbsp;% across all treatments (Fig.&nbsp;\@ref(fig:fig-1)e, \@ref(fig:fig-2)e).
As with the tiller incidence and severity, the effects of both the N100 and N120 treatments on leaf severity resulted in higher values and were and clearly different than the N0 treatment (Fig.&nbsp;\@ref(fig:fig-4)e, \@ref(fig:fig-5)e).
The AWD water management treatment was again undecided (Fig.&nbsp;\@ref(fig:fig-3)e, \@ref(fig:fig-7)a).

@@

The changes to the inoculation methodology resulted in a higher rate of infection in 2016, with the N180 treatment reaching a maximum value of 98&nbsp;% incidence at the third observation (Fig.&nbsp;\@ref(fig:fig-2)b).
The N treatment N180 value was higher and clearly different than the control N60 treatment (Fig.&nbsp;\@ref(fig:fig-4)b, \@ref(fig:fig-5)b). As in the 2015 study, water management treatment was undecided (Fig.&nbsp;\@ref(fig:fig-3)b, \@ref(fig:fig-5)b).

As with the tiller incidence, the tiller severity increased with the changed inoculation methods with a maximum of 7.6% for the puddled treatment (Fig.&nbsp;\@ref(fig:fig-2)d).
In 2016 the N180 treatment was again higher and clearly different from the N60 treatment, with N180 severity being higher (Fig.&nbsp;\@ref(fig:fig-4)d, \@ref(fig:fig-5)d).
The AWD water management, which was lower than the puddled treatment, was also clearly different (Fig.&nbsp;\@ref(fig:fig-3)d, \@ref(fig:fig-6)b).

In 2016 the higher N rate, N180, were clearly different, higher, than the control treatments, N60. However the water management treatment was undecided (Fig.&nbsp;\@ref(fig:fig-3)f, \@ref(fig:fig-5)b, \@ref(fig:fig-4)b, \@ref(fig:fig-7)b).

@@

x <- flextable(x)
x <- fontsize(x, part = "header", size = 10)
x <- font(x, part = "header", fontname = "Times")
x <- font(x, fontname = "Times")
x <- fontsize(x, size = 10)
x <- set_caption(x, 
  "(\\#tab:Table-2) An 89 % test for practical equivalence for 2015 experiment. Response variables are shown as T. Inc., Tiller Incidence; T. Sev., Tiller Severity; L. Sev., Leaf Severity. The parameters are WMGTAWD, irrigation treatment AWD; NRTEN100, nitrogen rate of 100 k/ha; NRTEN120 kg/ha. ROPE low and ROPE high provide the limits of the Region of Practical Equivalence (ROPE). ROPE Percentage is the proportion of the Highest Density Interval (HDI) that lies inside the ROPE. All points within this interval have a higher probability of density than points outside the interval. ROPE Equivalence is the test result, either \"rejected\", \"accepted\" or \"undecided\". \"HDI Low\" and \"HDI High\" are the lower and upper HDI limits for the parameters."
  )
x
```

@@

x <- flextable(x)
x <- fontsize(x, part = "header", size = 10)
x <- font(x, part = "header", fontname = "Times")
x <- font(x, fontname = "Times")
x <- fontsize(x, size = 10)
x <- set_caption(x, 
  "(\\#tab:Table-3) An 89 % test for practical equivalence for 2015 experiment. Response variables are shown as T. Inc., Tiller Incidence; T. Sev., Tiller Severity; L. Sev., Leaf Severity. The parameters are WMGTAWD, irrigation treatment AWD; and NRTEN180, nitrogen rate of 180 k/ha. ROPE low and ROPE high provide the limits of the Region of Practical Equivalence (ROPE). ROPE Percentage is the proportion of the Highest Density Interval (HDI) that lies inside the ROPE. All points within this interval have a higher probability of density than points outside the interval. ROPE Equivalence is the test result, either \"rejected\", \"accepted\" or \"undecided\". \"HDI Low\" and \"HDI High\" are the lower and upper HDI limits for the parameters.")
x
```
